Stability and Chaos in Multi-agent Reinforcement Learning

### 2020-10-09 aamas2021-formatting-instructions/aamas21-formatting-instructions/sample.tex
minor mods

We note that the numerical experiments confirm the predictions suggested by the analytic results shown in Figure \ref{sec::Theory}. We notice that convergence occurs almost only for low values of $\alpha$. This observation remains as we increase the value of $N$. As noted, the analytic result over-estimates the region of instability due to the assumptions of $N \rightarrow \infty$ and $\frac{1}{p-1} << 2$ made during the derivation. Remaining discrepancies are partially due to the aforementioned approximation in the calculation of $q$, $\chi$ and the fact that the analytic result considers a continuous time approximation of the expected behaviour of Q-Learning. We expect that testing over a greater number of payoff realisations and initial conditions will yield a more representative assessment of the average behaviour of Q-Learning. However, running these experiments is a computationally expensive procedure: for $p$ players and $N$ actions we require operations on $p$ matrices with $N^{p}$ elements. As such, due to a reduced availability of computational facilities, a large scale averaging was not possible.

However, due to the abundance of non-convergent behaviour, shown by the dark regions in Figure \ref{fig:NumericalExperiments} other than at low values of $\alpha, \tau$, we can conclude that the implications 1, 2 are verified. Indeed implications 3 and 4 are also verified as the proportion of games converged decrease monotonically with $\alpha, \tau$, but there does not appear to be any discernible dependence on $\Gamma$. Finally, we can conclude that implication 5 is verified by comparing Figures \ref{fig:pvariation} and \ref{fig:pVarianceExpt}. It is clear that the instability of the system increases monotonically with $p$. A point to note is that Figure \ref{fig:pVarianceExpt} shows a logarithmic dependence of the instability on $p$, whilst Figure \ref{fig:pvariation} suggests that the dependence should be exponential. Due to the limited range of $p$ that was used to generate Figure \ref{fig:pVarianceExpt}, it cannot yet be concluded that the dependence on $p$ is logarithmic. However, the important result is that the unstable behaviour increases, regardless of parameter choice as $p$ increases, which is corroborated by both theory and experiment. This phenomenon was referred to by Sanders et. al as `the prevalence of chaotic dynamics in games with many players' \cite{Sanders2018}.

A point which we wish to discuss here is the difference in the result found w.r.t.~\cite{Sanders2018}, which considered the stability of \textit{Experience Weighted Attraction} (EWA). Namely, Sanders et al.~found that convergence is seen for higher values of $\alpha$, whereas lower values give rise to chaos, the opposite of what is found here. The reason for this can be seen in the update equation of Q-Learning (\ref{eqn::Qupdate}), whereby
smaller values of $\alpha$ result in the agent placing a lower weight on the reward received at each step. As such, lower values of $\alpha$ result in the agent taking more conservative steps and yields a higher probability of convergence. In contrast, the update for EWA does not discount the reward received at all and, instead, only discounts the previous knowledge of the Q-value based on higher choices of $\alpha$. In essence, the difference is the fact that Q-Learning discounts new information by a factor of $\alpha$, whilst EWA does not. Yet, the difference in stability caused by this change is significant. This highlights the importance of performing analyses such as the present work; it allows for a method to analytically compare the difference between learning algorithms and, for practitioners, ensure that the appropriate algorithm is chosen for the parameters of their specific task.

>>>>

We note that the numerical experiments confirm the predictions
suggested by the analytic results in Figure
\ref{sec::Theory}. We notice that convergence occurs almost only for
low values of $\alpha$. This observation remains as we increase the
value of $N$. As noted, the analytic result over-estimates the region
of instability due to the assumptions of $N \rightarrow \infty$ and
$\frac{1}{p-1} << 2$ made during the derivation. Remaining
discrepancies are partially due to the aforementioned approximation in
the calculation of $q$, $\chi$ and the fact that the analytic result
considers a continuous time approximation of the expected behaviour of
Q-Learning. We expect that testing over a greater number of payoff
realisations and initial conditions will yield a more representative
assessment of the average behaviour of Q-Learning. However, running
these experiments is a computationally expensive procedure: for $p$
players and $N$ actions we require operations on $p$ matrices with
$N^{p}$ elements. As such, due to a reduced availability of
computational facilities, a large scale averaging was not possible.

However, due to the prevalence of non-convergent behaviour, shown by
the dark regions in Figure \ref{fig:NumericalExperiments} other than
at low values of $\alpha, \tau$, we conclude that implications (1),
(2) are verified. Indeed, implication (3) and (4) are also verified as
the proportion of games converged decrease monotonically with
$\alpha$, $\tau$, but there does not appear to be any discernible
dependence on $\Gamma$. Finally, we can conclude that implication 5 is
verified by comparing Figures \ref{fig:pvariation} and
\ref{fig:pVarianceExpt}. It is clear that the instability of the
system increases monotonically with $p$. A point to note is that
Figure \ref{fig:pVarianceExpt} shows a logarithmic dependence of the
instability on $p$, whilst Figure \ref{fig:pvariation} suggests that
the dependence should be exponential. Due to the limited range of $p$
that was used to generate Figure \ref{fig:pVarianceExpt}, it cannot
yet be concluded that the dependence on $p$ is logarithmic. However,
the key point is that, as $p$ increases, the unstable behaviour
increases, regardless of parameter choice, which is corroborated by
both theory and experiments. This phenomenon was referred to in
\cite{Sanders2018}
%by Sanders et. al
as `the prevalence of chaotic dynamics in games with many
players'.

A point which we wish to discuss here is the difference in the result
found w.r.t.~\cite{Sanders2018}, which considered the stability of
\textit{experience weighted attraction} (EWA). Namely, Sanders et
al.~found that convergence is seen for higher values of $\alpha$,
whereas lower values give rise to chaos, the opposite of what is found
here. The reason for this can be seen in the update equation of
Q-Learning (\ref{eqn::Qupdate}), whereby smaller values of $\alpha$
result in the agent placing a lower weight on the reward received at
each step. As such, lower values of $\alpha$ result in the agent
taking more conservative steps and yields a higher probability of
convergence. In contrast, the update for EWA does not discount the
reward received at all and, instead, only discounts the previous
knowledge of the Q-value based on higher choices of $\alpha$. In
essence, the difference is the fact that Q-Learning discounts new
information by a factor of $\alpha$, whilst EWA does not. Yet, the
difference in stability caused by this change is significant. This
highlights the importance of performing analyses such as the present
work; it allows for a method to analytically compare the difference
between learning algorithms and, for practitioners, ensure that the
appropriate algorithm is chosen for the parameters of their specific
task.
